---
title: "GV300 Showcase: Electoral Participation in the UK"
output: distill::distill_article
---

## Introduction

This page showcases a piece of work from my GV300 assignment on electoral participation in the United Kingdom. The aim of the project was to explore whether all votes are equally represented and to analyse differences in constituency sizes and voter turnout across the UK. Using R and the tidyverse packages, I cleaned election data and created visualisations to better understand patterns in electoral participation.

## Example Code

Below is an example of the code used to clean and prepare the dataset for analysis.

```{r}
library(tidyverse)

uk_csv <- "https://electionresults.parliament.uk/general-elections/6/candidacies.csv"
uk_raw <- readr::read_csv(uk_csv, show_col_types = FALSE)

names(uk_raw) <- names(uk_raw) %>% 
  tolower() %>% 
  str_replace_all("\\s+", "_")

uk1 <- uk_raw %>%
  filter(!`election_is_by-election`)

uk2 <- uk1 %>%
  select(
    constituency = constituency_name,
    country = country_name,
    valid_votes = candidate_vote_count,
    electorate = electorate
  )
```

```{r}
ggplot(uk, aes(electorate)) +
  geom_histogram(binwidth = 5000, color = "white") +
  labs(
    title = "Distribution of Constituency Electorates in the UK",
    x = "Electorate size",
    y = "Number of constituencies"
  ) +
  theme_minimal()
``` 
### Explanation
The code above loads election data from the UK Parliament dataset and prepares it for analysis. It removes by-elections, keeps only the key variables, combines candidate-level rows into constituency-level totals, and calculates turnout as a percentage of the electorate. This makes the dataset suitable for comparing constituency size and turnout across the UK.
The visualisation shows the distribution of electorate sizes across UK constituencies. Most constituencies are clustered around roughly 70,000 to 78,000 voters, but there are also some noticeably smaller and larger constituencies. This matters because each constituency elects one Member of Parliament regardless of its size, which means that votes may not carry exactly the same weight everywhere.